---
title: "STAT 3350 Assignment 1, Winter 2020 - due Wednesday, Jan 15, at beginning of class"
author: "Chaojun Ma"
date: 'Banner:  B00726514'
output:
  pdf_document: default
  html_document: default
  word_document: default
---

1. A computer program has produced the following output for a hypothesis testing problem:

\begin{verbatim}
Difference in sample means: 2.35
Degrees of freedom: 18
Standard error of the difference in sample means: ?
Test statistic: t0 = 2.01
P‐value: 0.0298
\end{verbatim}

   + 1a. What is the value of $s_p$, the pooled estimate of $\sigma$.
```{r}
#do your calculations in R here
SE=2.35/2.01;
SP=SE/sqrt(1/5);SP
cat("The SP value is ",SP)
```

   + 1b. Is this a two‐sided or a one‐sided test?
```{r}
pvalue1b=1-pt(2.01,18);pvalue1b
cat("The result same as above output, since is one sided test")
```
   
just type in your answer here.

   + 1c. If $\alpha = 0.05$,  what is your conclusion?
```{r}
cat("P value is smaller than 0.05, reject null hypothesis at alpha = 0.05")
```
   
 reject the null hypothesis, or do not reject the null hypothesis.
 
   + 1d. Find a 90 percent two‐sided CI on the difference in means.
```{r}
#R calculations here
tobsvalue=abs(qt(0.1/2,18));tobsvalue
LowerBound=2.35-tobsvalue*SE;
UpperBound=2.35+tobsvalue*SE;
cat("Since the lowerbound value is ",LowerBound,"the upperbound value is ",UpperBound)
```


2.15 Consider the computer output shown below.

\begin{verbatim}

Two‐sample T for Y1 vs Y2
         N     Mean     Std. Dev.    SE Mean
Y1      20    50.19       1.71          0.38
Y2      20    52.52       2.48          0.55
\end{verbatim}


When testing for the difference  $\mu_1 - \mu_2$:

\begin{verbatim}
estimated difference: -2.33341
95% CI for difference: (-3.69547, -0.97135)
T‐Test of difference = 0 (vs not =): T‐Value = -3.47
P‐Value = 0.001 DF = 38
Both use Pooled Std. Dev. = 2.1277
\end{verbatim}

   + 2a. Can the null hypothesis be rejected at the 0.05 level? Why?
```{r}
cat("P value is 0.001, P value smaller than critical value at 0.05 level, rejceted null hypothesis")
```

   + 2b. Is this a one‐sided or a two‐sided test?
```{r}
pvalue=pt(-3.47,38);pvalue
pvalue2=2*(1-pt(abs(-3.47),38));pvalue2
cat("since one-sided value is 0.000655 not equal to above result however two-sided value is 0.001311611 is equal to above")
cat(". Therfore this is a two-sided test")
```

   + 2c.  If the hypotheses had been $H_0: \mu_1-\mu_2 = 2$ versus $H_1: \mu_1-\mu_2 \ne 2$,
would you reject the null hypothesis at the 0.05 level?,
```{r}
cat("At 0.05 level of the 95% confidence interval does not contain 2 then reject H0")
```

   + 2d.  $H_0: \mu_1-\mu_2 = 2$ versus $H_1: \mu_1-\mu_2 < 2$, would you reject the null hypothesis
   at the 0.05 level? Can you answer this question without doing any additional calculations? Why?
```{r}
cat("At 0.05 level of the 95% confidence interval does not contain 2 then reject H0")
```

   + 2e. Use the output and the R $pt$ command to find a 95 percent upper confidence bound on the difference in means.
```{r}
tvalue = qt(0.95,38)
upperbound = -2.33341+tvalue*2.1277*sqrt(1/10);upperbound

```

As an example, to find the upper 10'th percentile of the $t$ distribution with 12 degrees of freedom, you can
use the following syntax.

```{r}
pct=pt(.9,12) 
pct
```

   + 2f. What is the P-value if the hypotheses are 
   
   $H_0: \mu_1-\mu_2 = 2$ versus $H_1: \mu_1-\mu_2 \ne 2$?


```{r}
TS=  (-2.33341-2)/(2.1277*sqrt(1/10));TS
pvalue = 2*(1-pt(abs(TS),38))
pvalue
cat("The pvalue is ",pvalue)
```


3.  A comparative experiment is to be carried out to compare two populations means with the hypotheses
$H_0: \mu_1 - \mu_2 = 0$ vs $H_0: \mu_1 - \mu_2 > 0$. Assuming that the common
standard deviation of the populations is $2$, that testing will be carried out at level $\alpha=.05$, and that
equal sample sizes will be used, what is the sample size PER GROUP that is required:

   +  eg. to achieve 80% power when $\mu_1-\mu_2 = 2$
```{r}
ss=2*(qnorm(1-.05/2)-qnorm(1-.8))^2/((2/2)^2)
ss
ceiling(ss) #round ss up to the next integer
```
   
   + 3a.   to achieve 90% power when $\mu_1-\mu_2 = 2$ 
```{r}
ssResulta=2*(qnorm(0.05)+qnorm(0.1))^2/(2/2)^2;ssResulta
cat("The samle size is ",ceiling(ssResulta))
```
   
   
   + 3b.   to achieve 80% power when: $\mu_1-\mu_2 = 1$
```{r}
ssResultb=2*(qnorm(0.05)+qnorm(0.2))^2/(1/2)^2;ssResultb
cat("The samle size is ",ceiling(ssResultb))

```
   
   
   + 3c.   to achieve 80% power for a 2 sided test with alternative $H_1: \mu_1 - \mu_2 > 0$ at level $\alpha=.05$ when $\mu_1-\mu_2 = 1$
```{r}
ssResultc=2*(qnorm(0.05/2)+qnorm(0.2))^2/(1/2)^2;ssResultc
cat("The samle size is ",ceiling(ssResultc))

```
   
   + 3d.   to achieve 80% power when $\mu_1-\mu_2 = 2$ when testing is at level $\alpha=.01$
```{r}
ssResultd=2*(qnorm(0.01)+qnorm(0.2))^2/(2/2)^2;ssResultd
cat("The samle size is ",ceiling(ssResultd))

```
   
   + 3e.   to achieve 80% power when $\mu_1-\mu_2 = 2$ when testing is at level $\alpha=.10$.   
```{r}
ssResulte=2*(qnorm(0.10)+qnorm(0.2))^2/(2/2)^2;ssResulte
cat("The samle size is ",ceiling(ssResulte))

```
   

4.  A comparative experiment is to be carried out to compare two populations means with the hypotheses
$H_0: \mu_1 - \mu_2 = 0$ vs $H_0: \mu_1 - \mu_2 > 0$. Assuming that the common
standard deviation of the populations is $2$, that testing will be carried out at level $\alpha=.05$.  What is
the power to reject the null hypothesis when equal sample sizes of 10 are used in EACH group and $\mu_1-\mu_2 = 1$.
```{r}
powerResult4=1-pnorm(qnorm(0.95)-1/2*sqrt(5));powerResult4
cat("The power to reject null hypothesis is ",powerResult4)
```
