---
title: "STAT 3350 Assignment 2, Winter 2020 - due Wednesday, Jan 29, at beginning of class"
author: "Chaojun Ma"
date: 'Banner:  B00726514'
output:
  pdf_document: default
  word_document: default
  html_document: default
---

1. A computer ANOVA output is shown below. Fill in the missing entries.

+ 1a.Fill in the missing entries

````markdown
One‐way ANOVA
 
Source   DF        SS      MS   F   
Factor    4    987.71   246.93  33.095   
Error    25    186.53   7.4612
Total    29   1174.24 

````
+ 1b.  Calculate the p-value using the pf command in R.

```{r}

MSE = 186.53/25;
Fvalue = 246.93/MSE;
pvalue = 1-pf(Fvalue,4,25);#get Factor df 

cat("The P-avlue is:",pvalue);

```

2. The tensile strength of Portland cement is being studied. Four
different mixing techniques can be used economically. A completely
randomized experiment was conducted, with four measurements of tensile
strength made for each of 4 mixing techniques, leading to the
following data.

````markdown
Mixing technique     Tensile Strength
    A                 3129 3000 2865 2890
    B                 3200 3300 2975 3150
    C                 2800 2900 2985 3050
    D                 2600 2700 2600 2765
````

   + 2a.  Make a boxplot and point plot to examine the distribution of
   tensile strength for different mixing techniques.  Does it look as if
   there are differences in the distributions of tensile strength?

```{r}
#put your R commands here

Rawdata=c(   3129, 3000, 2865, 2890,
             3200, 3300, 2975, 3150,
             2800, 2900, 2985, 3050,
             2600, 2700, 2600, 2765) #y value 
treatment=c(rep(1,4),rep(2,4),rep(3,4),rep(4,4));#make a matrix 4*4 such like 1 1 1 1
                                                 #                            2 2 2 2 or you can using c= (1,1,1,1...) 
                                                 # x value 

boxplot(Rawdata~treatment,xlab = "Mixing Tech", ylab = "Tensile Strength")

plot(treatment,Rawdata,xlab = "Mixing Tech", ylab = "Tensile Strength") # or y~x 

```

```{r}
cat("According to above figure, we can get following conclusion: The tensile strength is differnted by differnt mixing techniques");
```
   
   + 2b. Use the tapply command to calculate the sample mean tensile strength
   for each mixing technique.

```{r}
#put your R commands here

means=tapply(Rawdata, treatment, FUN = mean);
means

```
   + 2c. Carry out an analysis of variance using the lm and anova commands
   in R.  Show the summary output for the lm fit, and the anova output.
   What is the p-value for the ANOVA?  What is your conclusion when testing
   at level $\alpha=.05$.

```{r}
#put your R commands here

tfactor = as.factor(treatment)
model = lm(Rawdata~tfactor)
summary(model)
anova(model)

```
```{r}
cat("As above result shows, the p-value is 0.000487. The alpha =.05 and p-value less than alpha. Therefore,reject the null hypothesis, at least one group of mean are not eqaul ")
```

   + 2d.  Calculate the sums of squares for the following 3 orthonormal
   contrasts:

```{r}
C1= c(1,1,-1,-1)/sqrt(4)
C2= c(1,-1,1,-1)/sqrt(4)
C3= c(1,-1,-1,1)/sqrt(4)# When there are a treatments, a−1 orthogonal contrasts partition the treatment sum of squares (having                            a−1 degrees of freedom) (refence:STAT 3350)

SS1 = 4*sum(C1*means)^2;SS1# formula n*sum (c*means)^2    >>CH3 
SS2 = 4*sum(C2*means)^2;SS2
SS3 = 4*sum(C3*means)^2;SS3

#check SS1+SS2+SS3 = SST factor  directly copy from lecture note ©️STAT 3350
print(paste("sum of the contrast sum of squares is ",SS1+SS2+SS3))
```

   + 2e.  Use Tukey's method with $\alpha=.05$ to make confidence intervals
   for all pairwise differences of population means.

```{r}
#put your R commands here

MSE = 12826
critValue=qtukey(1-0.05,4,12);

means[1]-means[2]+c(-1,1)*critValue*sqrt(MSE/4);
means[1]-means[3]+c(-1,1)*critValue*sqrt(MSE/4);
means[1]-means[4]+c(-1,1)*critValue*sqrt(MSE/4);#U1 not equal to U4

means[2]-means[3]+c(-1,1)*critValue*sqrt(MSE/4);
means[2]-means[4]+c(-1,1)*critValue*sqrt(MSE/4);#U2 not equal to U4

means[3]-means[4]+c(-1,1)*critValue*sqrt(MSE/4);#U3 not equal to U4

cat("Above results show U1 not equal to U4, U2 not equal to U4, U3 not equual to U4. It sugguests that mixing techniques D is differ mixing techniques A, B and C.");


```

    What are your conclusions in terms of which mixing techniques give
    different mean tensile strengths.

3.  The response time in milliseconds was determined for three
different types of circuits that could be used in an automatic valve
shutoff mechanism. The results from a completely randomized experiment
are shown in the following table:

````markdown
Circuit        Response
 type            time
I             9 12 10  8 15
II           20 21 23 17 30
III           6  5  8 16  7
````

   Construct a pair of orthogonal, normalized, contrasts assuming
   that at the outset of  the experiment you suspect that the response
   time of circuit type 2 is different from the other two circuit types.
   Construct 95% confidence intervals for the two contrasts.

```{r}
#enter your R commands here n=5 a=3

y=c(9,12,10,8,15,
   20,21,23,17,30,
   6,5,8,16,7);

treatment3=c(rep("A",5),rep("B",5),rep("C",5));

means3=tapply(y, treatment3, FUN = mean);
means3

tfactor = as.factor(treatment3)
model = lm(y~tfactor)
summary(model)
anova(model)


```
```{r}
MSE=16.9;
critValue=qt(0.025,12);

C1 = c(1,-2,1)/sqrt(6);
C2 = c(1,0,-1)/sqrt(2);

CI1 = (means3[1]-2*means3[2]+means3[3])/sqrt(6)+c(-1,1)*critValue*sqrt(MSE/5);
CI1

CI2 = (means3[1]-means3[3])/sqrt(2)+c(-1,1)*critValue*sqrt(MSE/5);
CI2

```


```{r}
samplea <- c(1:50,0)
```


