---
title: "STAT 3350 Assignment 4, Winter 2020 - due Friday, Feb 28, at beginning of class"
author: "Chaojun Ma"
date: 'Banner:  B00726514'
output:
  pdf_document: default
  word_document: default
  html_document: default
---


1. A study used the $^{13}C-UBT$ test on a group of patients infected with
H. pylori, and on another group who were not infected, with the
following results, which shows the measurements on the outcome variance $\Delta^{13}CO_2$.

\begin{table}[h]
\center
\begin{tabular}{rr}
\multicolumn{2}{c}{Infected}\\
Yes&No\\
\hline
4.8&2.9\\
4.3&3.6\\
3.9&3.8\\
3.7&\\
\hline
\end{tabular}
\end{table}

Using a permutation test, assess whether there is difference in mean $\Delta^{13}CO_2$ between the two
groups using a permutation test. Use the hypotheses:

$H_0: \mu_{Yes} = \mu_{No}$ vs 
$H_1: \mu_{Yes} > \mu_{No}$

(1a) If the difference of sample means $\bar X_{yes} - \bar X_{no}$ is used as the test statistic, what is the observed value of the test statistic? 
```{r}
cat("The result value is 0.742");
```


   (1b)  How many assignments of subjects to treatments are there with 4 subjects in the first group and 3 subjects in the second group? 

```{r}
result1b = choose(7,4);
cat("Reuslt is",result1b);
```


   (1c) How many assignments of subjects to treatments are there leading to a value for the test statistic which is at least as contrary to the null hypothesis as was the
original data? 

```{r}
cat("The answer is two assignment ");
```


   (1d)  What is the p-value? 
```{r}
cat("The palue is" , 2/35);
```


2. The following code generates observations from a two factor
analysis of variance model with two levels of factor A, three
levels of factor B, and 5 replicates, and plots the sample means of $y$ vs the levels of factor B, with separate plots for the two levels of factor A.

```{r}
set.seed(42);
y=c(rnorm(5,5),rnorm(5,10),rnorm(5,10),rnorm(5,10),rnorm(5,5),rnorm(5,5))

A=as.factor(c(rep(1,15),rep(2,15)));
B=as.factor(c(rep(1,5),rep(2,5),rep(3,5),rep(1,5),rep(2,5),rep(3,5)))

anova(lm(y~A*B))

interaction.plot(B,A,y);
```

Use the anova and lm commands to carry out a two way analysis of
variance.  Using level $\alpha=.05$, which of the effects (main effect of A, main effect of B, or interaction) are significant, and
what are the assoicated p-values? 

```{r}
lm.out = lm(y~A*B);
anova(lm.out);
```

If you did the analysis correctly, you should find that the interaction and the main effect of A are significant, but the main effect of B is not significant, and that the SSE for this two way model
is 38.195.

The following calculates the sample mean of $y$ for each combination of A and B.

```{r}
tapply(y,list(A,B),FUN=mean)
```

Use Tukey's procedure to determine which of the 6 means are
different.  As 6 means are being compared, you should use the percentile given by qtukey(.95,6,24), as there are 6 means
being compared, and 24 degrees of freedom associated with the
MSE. 
(Hint: Using Tukey's procedure, for any two sample means which differ by
more than  $qtukey(1-\alpha,a,df) \sqrt{MSE/n}$, the associated population means are declared to be different.  Hence you only
need to calculate this value, and then look at the differences in
means in the above table.  You don't need to calculate all of 
the 6 intervals.)

   (2a) What is the value of $K=qtukey(1-\alpha,a,df) \sqrt{MSE/n}$?
   
```{r}
K=qtukey(0.95,6,24) *sqrt(1.591/30); K
```
   
   (2b) Using this value, which of the 6 population means are determined to be different using Tukey's method.  For example, the population means for (A=1,B=1) and (A=1,B=2) are declared to be different if
you have calculated the correct value of $K$.  Just list the pairs
which are declared different.

```{r}
cat("Following pairs are declared different:
    A1B1-A1B2;A1B1-A1B3;A1B1-A2B1
    A1B2-A2B1;A1B2-A2B2;A1B2-A2B3
    A1B3-A2B1;A1B3-A2B2;A1B3-A2B3
    A2B1-A2B2;A2B1-A2B3");
```

The width of the 95% Tukey confidence intervals is $2 qtukey(.95,6,24) \sqrt{MSE/n}$, with the MSE taken from the ANOVA table.

When comparing "6 choose 2"=15 pairs of means, the Bonerroni corrected 95\% $t$ confidence intervals for $\mu_i - \mu_j$
  are of the form

$$\bar y_{i.} - \bar y_{j.} \pm \sqrt{2} qt(1-.025/15,df) \sqrt{MSE/n}$$
where $df$ is the degrees of freedom associated with the MSE in the ANOVA table, and the multiplier $\sqrt{2}$ is needed because the contrast $\mu_i-\mu_j$ is not normalized.  The width of this interval is
$2 \sqrt{2}  qt(1-.025/15,df) \sqrt{MSE/n}$.


  (2c) Which of the intervals (Tukey of Bonerroni corrected $t$) is wider?  (You needn't look at the common factor $\sqrt{MSE/n}$, just the 
  multipliers $2 qtukey$ and $2 \sqrt{2} qt$.)
```{r}
qtukey=qtukey(0.95,6,24);qtukey
tcrit=sqrt(2)*qt(1-0.025/15,24);tcrit

```

   (2d) Both Tukey and Bonerroni corrected $t$ intervals control the familywise Type I error rate to be at most 5\%.  Which of
   the intervals do you prefer, and why?

```{r}
cat("I perfer the Tukey's method, because Turkey's method can modify to exactly idea alpha, moreover, Turkey's method can provide shorter confidence intervals. ")

```

