---
title: "STAT 3340 Assignment 1, Fall 2019 - due Thursday, Sept 26, at beginning of class"
author: "Chaojun Ma"
date: 'Banner:  B00726514'
output:
  html_document: default
  pdf_document: default
  word_document: default
---
=================================================================================================================

You should process this Rmd file in R markdown, after adding the appropriate 
R code or text where appropriate.  You should add your name and banner
number in the appropriate place above.  Then "Knit" to pdf, Word or html
to produce your output document.  Most of you will NOT be able to produce
a pdf document because your computer will not have the requisite software,
so you should use Word or html output, whichever looks better.  Once you
are happy with your answers, print a hard copy, which is what is to be
submitted for grading.


1. Find the equation of the line which passes through the points (1,1) and (4,5).

Answer: 

$$y=-\frac{1}{3} + \frac{4}{3} x$$

2. Suppose you are given three data points (1,4), (2,6) and (3,7) and the line $y = 1 + 4x$. 
Give the three residuals and their sum of squares.

    $r1=4-5=-1$ 
    
    $r2=6-9=-3$
    
    $r3=7-13=-6$
    
    $$\sum_{i=1}^3 r_i^2 =1+9+36=46 $$

3. Some data gives the summaries: $n = 10$, $\sum_{i=1}^{10}  x_iy_i = 100$, $\sum_{i=1}^{10} x_i = 20$ and $\sum_{i=1}^{10} y_i = 10$. 
   Suppose that the response $y$ is temperature in degrees Celcius. 

    + 3.a) $S_{xy}$=
```{r}
#Answer:
#1
#2
#3
#4
#5
#6




```
    
    
    + 3.b) If the response was converted to temperature in degrees Fahrenheit $y'$, so that $y' = 32
+1.8y$, what is $\sum_{i=1}^{10} y'_i$? (this points out how the sum operation
works when applied to a linear transformation of the original variable)

```{r}

#Answer:
#1
#2
#3
#4
#5
#6



```

    + 3.c) If the response was converted to temperature in degrees Fahrenheit $y'$, so that $y' = 32+1.8y$, what is $S_{xy'}$? (this illustrates how the
sum of produces operation works when one of the variables is replaced by
a linear transformation of the original variable)

```{r}




#Answer:
#1
#2
#3
#4
#5
#6

```


4. A random sample of 11 elementary school students is selected,
  and each student is measured on a creativity score (x) using a
  well-defined testing instrument and on a task score (y) using a new
  instrument. The task score is the mean time taken to perform several
  hand-eye coordination tasks. The data are as follows.
  
```{r}
 x=c(35,37,50,69,84,40,29,42,51,45,31)
 y=c(3.9,3.9,6.1,4.3,8.8,2.1,5.7,3.0,7.1,7.3,3.3)
 plot(x,y,xlab="Creativity score",ylab="Task score");
 
```
 
 Use R to do the following questions. You just need to include the appropriate R commands in the markdown document.
 
   + 4a) Calculate the summaries $S_{xy}$, $S_{yy}$ and $\bar{Y}$. R commands to calculate  $S_{xx}$ and $\bar{X}$ have been given in order to show you how to include the R syntax in your markdown document.
  
 
```{r}

N=length(x)
xbar=sum(x)/N ;  cat("xbar = ", xbar)
Sxx=sum((x-xbar)^2) ; cat("Sxx = ",Sxx)
ybar=sum(y)/N ; cat("ybar = ", ybar);
Sxy=sum((x-xbar)*(y-ybar)); cat("Sxy = ", Sxy);
Syy=sum((y-ybar)^2); cat("Syy = ", Syy);

#IN THE SAME WAY, YOU WILL NEED TO DROP THE # IN SEVERAL SUBSEQUENT QUESTIONS
```

   + 4b) Use these summaries to  calculate the least squares estimates of the  intercept and slope.
  
```{r}

 beta1 = Sxy/Sxx ; cat("beta1 =", beta1) ;
 beta0=ybar- beta1 * xbar ;cat("\nbeta0 =", beta0)  

```
  
   + 4c) Calculate the predicted (fitted) values and the residuals.
   
```{r}
fit=beta0+beta1*x;
resids=y-fit;
fit;
resids
```
   +  4d) Plot the residuals (y axis) vs fitted values (x axis).


```{r}
plot(fit,resids,xlab = "Fitted values",ylab="Rediduals");
```

   + 4e) Do the residuals look random? If not, what does the pattern of the residuals suggest?  (NO GRADE FOR THIS QUESTION.  IT'S REALLY HARD TO ASSESS
   RANDOMNESS VISUALLY WITH ONLY A FEW DATA POINTS)
   
   
   + 4f) Calculate the mean of the residuals
to verify it is zero, and the correlation of the residuals with $X$ to verify it is also zero.  The R command to calculate
the correlation between $u$ and $v$ is cor(u,v).  NOTE THAT THESE ARE 0 ONLY
TO MACHINE PRECISION.  THE COMPUTER DOESN'T DO EXACT ARITHMETIC, SO EVEN THOUGH WE KNOW THAT THESE QUANTITIES ARE 0 IN THEORY, THE CALCULATED VALUES 
WILL BE VERY CLOSE TO 0, BUT NOT EXACTLY 0.  NOTE, FOR EXAMPLE, THAT
1.4e-3 means 1.4x10^(-3), or .0014.  THIS IS TYPICALLY HOW MOST COMPUTER
PROGRAMS REPRESENT SCIENTIFIC NOTATION.

```{r}
mean(resids);
cor(resids,x);

```

  + 4g) Calculate and print the error, regression and total sums of squares, using the data summaries.

```{r}
SST=Syy; SST
#SSE=      sum of (residuals-mean(residuals))^2
# NOTE THAT THE MEAN OF THE RESIDUALS IS EQUAL TO 0, SO SSE IS TYPICALLY CALCULATED AS
SSE=sum(resids^2)；SSE
SSR=SST - SSE;SSR
```


5. The length of a species of fish is to be represented as a function
of the age and water temperature.  The fish are kept in tanks
at 25, 27, 29 and 31 degrees Celsius.  After birth, a test specimen
is chosen at random every 14 days and its length measured.  The data can be entered into R as

```{r}
fish=read.csv("http://bsmith.mathstat.dal.ca/stat3340/Data/fish.csv",header=T)
pairs(fish)
```

The pairwise plots don't really give an adequate picture of what is going on.
The following plot of length vs age, with different temperatures identified
suggests that growth is impeded by high temperatures.

```{r}

  color=rep("black",dim(fish)[1])
  color=ifelse(fish$temp==27,"blue", color)
  color=ifelse(fish$temp==29,"green",color)
  color=ifelse(fish$temp==31,"brown",color)
  legend.txt=c("25 degrees","27 degrees","29 degrees","31 degrees")

plot(fish$age,fish$length,col=color,ylab="length",xlab="age",cex=2,pch="*")
legend(100,2000,legend=legend.txt,text.col=c("black","blue","green","brown"))

```

The multiple regression model $length = \beta_0+\beta_1 age +\beta_2 temp +\epsilon$
was fit using R, as follows.


```{r}
fish.out=lm(length~age+temp, data=fish)
summary(fish.out)
anova(fish.out)
```

  +  Test for the overall significance of the regression.  
    
The hypotheses are: $H_0: \beta_1 = \beta_2 = 0$ vs
$H_A:\mbox{ one or more of } \beta_1, \beta_2 \ne 0$.

  + 5a) The observed value of the $F$ statistic is 84.98
  
  + 5b) he numerator degrees of freedom is 2
  
  + 5c) The denominator degrees of freedom is 41
  
  + 5d) Calculate the  p-value
```{r}
 1-pf(84.98,2,41)  # and include the appropriate Fobs and degrees of freedom.
```



  + 5e) Use the lm command to fit a model which includes only the variable age.

```{r}
ageonly.out = lm (length~age, data=fish) # and include the appropriate model syntax
summary(ageonly.out);
```

Test the appropriate hypothesis concerning "temperature".

  + 5f) state the hypotheses:  $H_0:beta 2 = 0 $  vs  $H_A:beta 2 \ne 0$
    
  + 5g) Use the ANOVA command to compare the two models (here fish.out and ageonly.out)

```{r}
anova(fish.out, ageonly.out)
```

   + 5g(i)The observed value of the test statistic is 6.9202
   + 5g(ii) The numerator degrees of freedom is 1
   + 5g(iii)The denominator degrees of freedom is 41
   + 5g(iv) Calculate the p-value
    
```{r}
1-pf(6.9202,1,41);
```
    
    
  + 5h) The model 
     $$length = \beta_0+\beta_1 age +\beta_2 temp + \beta_3 age*temp +\epsilon$$
    includes the interaction between temperature and age.  The model
    is fit in R using "lm(length\~age+temp+age:temp)".  An equivalent
    model statement is "lm(length\~age*temp)".  Fit this model in R,
    and then use the "anova" command to carry out the F test of significance of the interaction.
    
```{r}
lm.interac=lm(length~age+temp+age:temp, data=fish)
anova(lm.interac,fish.out)
```
  + 5h(i) The null hypothesis is $H_0:\mbox{hypothesis about  a } \beta 3 =0$ parameter 
  + 5h(ii) The alternative hypothesis is $H_A: \mbox{hypothesis about a } \beta 3 \ne 0$ parameter
  + 5h(iii) The observed value of F is  3.8858
  + 5h(iv) The numerator degrees of freedom is 1
  + 5h(v) The denominator degrees of freedom is 40
  + 5h(vi) Calculate the p-value.

```{r}
1-pf(3.8858,1,40);
# Don't take the p-value from the above output.
# Rather, include your R Markdown code for the p-value calculation here.

```

6.  You can do this question by hand using basic ideas from linear 
algebra, or in R, remembering that you
can calulate the projection of y on x in R as the fitted values of
the regression of y on x through the origin, using 'fitted(lm(y~x-1))',
and you can get the residuals from that regression using
'residuals(lm(y~x-1))'.

   + 6a)  exercise 2.6.8 of Selinger, page 89.
```{r}
w=c(1,0,-2);
v=c(1,2,3);
fitted(lm(w~v-1));#model pass the origin no intercept 

```
   
   + 6b)  exercise 11.3.1 of Selinger, page 412.
```{r}
v11=c(1,2,3);
v12=c(2,6,0);
u11=v11;
u12=v12-fitted(lm(v12~u11-1))
u11;
u12;

```
   
   
   + 6c)  exercise 11.3.3 of Selinger, page 412.   
```{r}

v1=c(1,0,1,0);
v2=c(1,3,1,-1);
v3=c(2,4,2,2);
u1=v1;
u2=v2-fitted(lm(v2~u1-1));

u3=v3-fitted(lm(v3~u1-1))-fitted(lm(v3~u2-1));

u1;
u2;
u3;

```
   










