---
title: "STAT 3340 Assignment 4, Fall 2019 - due Tuesday, Nov 5, at beginning of class"
author: "Chaojun Ma"
date: 'Banner:  B00726514'
output:
  pdf_document: default
  html_document: default
  word_document: default
---
=================================================================================================================




Data were collected to relate the compressive strength of
concrete ($Y$) to a set of predictors
\begin{table}[h]
\center
\begin{tabular}{ll}
$X_1$&log(age)\\
$X_2$&cement\\
$X_3$&slag\\
$X_4$&fly.ash\\
$X_5$&superplasticizer\\
$X_6$&water\\
$X_7$&coarse.aggregate\\
$X_8$&fine.aggregate\\
\end{tabular}
\end{table}

\noindent
The partial regression output below shows the sequential sum of
squares as terms are added to the model.
\small
\begin{verbatim}

> cmt.out1=lm(Y~log(age)+cement+slag+fly.ash+superplasticizer+water+coarse.aggregate+
    fine.aggregate,data=cmnt)

> anova(cmt.out1)
Analysis of Variance Table

Response: Y
                  Df Sum Sq Mean Sq  F value    Pr(>F)    
log(age)           1 66.202  66.202 150.2657 < 2.2e-16 ***
cement             1 74.665  74.665 169.4744 < 2.2e-16 ***
slag               1 24.952  24.952  56.6354 5.652e-12 ***
fly.ash            1 26.017  26.017  59.0526 2.365e-12 ***
superplasticizer   1 14.837  14.837  33.6763 4.117e-08 ***
water              1 11.162  11.162  25.3365 1.445e-06 ***
coarse.aggregate   1  0.796   0.796   1.8063    0.1811    
fine.aggregate     1  0.350   0.350   0.7941    0.3744    
Residuals        141 62.120   0.441                       
---

\end{verbatim}

Assess whether $coarse.aggregate$ and 
$fine.aggregate$ are  linearly related to $Compressive~Strength$ once the 
other variables have been included in the model.

   + 1a) Evaluate the test statistic.
```{r}
F=((0.796+0.350)/2)/(62.120/141)
F
```
   
   + 1b) State the numerator and denominator degrees of freedom, and calculate the p-value 
using the pf command in R.

```{r}
pv=1-pf(F,2,141);
pv
```

2. The data set ``fish'' has data on fish lengths, age and
  water temperature.   

The data can be read from the .csv file, for example, as


```{r}
fish=read.csv("http://bsmith.mathstat.dal.ca/stat3340/Data/fish.csv",header=T)
```

   + 2a)
Using the {\em lm} command, fit the linear model
$$
y = \beta_0+\beta_1 x_1 + \beta_2 x_2 +\epsilon
$$ 
where $y$ is length, $x_1$ is $age$, and $x_2$ is $temperature$. 

Print the "summary" output for the model.

```{r}
lm.out=lm(length~age+temp,data=fish);
summary(lm.out);
```

   + 2b) Fit a quadratic model 
$$
y = \beta_0+\beta_1 x_1 + \beta_2 x_2 + \beta_3 x_1x_2 + \beta_4 x_1^2 + \beta_5 x_2^2 + \epsilon
$$ 

In order to enter a quadratic term, say $age^2$ the R syntax is 
$I(age\wedge2)$.

For example, a model for fish length which includes linear and quadratic
terms in age would be specified as 

$lm(length \sim age + I(age \wedge 2),data=fish)$

Print the "summary" output for the model.

```{r}
lm.quad= lm(length~age+temp+age*temp+I(age^2)+I(temp^2),data = fish)
summary(lm.quad)
```

   + 2c)  By comparing the error sums of squares in the summary outputs for the two models, test whether
the squared and product terms can be omitted from the quadratic model.


   + 2c(i) State the hypotheses.
```{r}
#H0:b3=b4=b5=0;
#Ha:at least one of them not equal to 0
```
   
   + 2c(ii) What is the observed value of the test statistic?
   
```{r}
anova(lm.out,lm.quad)
```
   
   + 2c(iii) Calculate the p-value using the $pf$ command in R.
```{r}
1-pf(74.254,3,38)
```

3.   In class we talked about how we can consider regression on $ X_1$ and $X_2$ to
be the result of three regressions. In this question we apply this approach to the $fish$
data, where we add the $age^2$ term to the model which already has the terms linear in $age$
and $temperature$.


   + 3a)
   Regress $length$ on $age$ and $temperature$. Include the $summary$ and $anova$ outputs.
   Store the residuals in $e_1$.
```{r}
lm1=lm(length~age+temp,data=fish)
summary(lm1)
anova(lm1)
e1=residuals(lm1)
```
   
   + 3b) Regress $age^2$ on $age$ and $temperature$. Store the residuals in $e_2$
```{r}
lm2=lm(I(age^2)~age+temp,data=fish)
e2=residuals(lm2)
```
   
   + 3c) Regress the residuals $e_1$ on the residuals $e_2$. Do not include an intercept. Use
the formula $lm(e1 \sim e2-1)$. Print the $anova$ output.
```{r}
lm3=lm(e1~e2-1)
summary(lm3)
anova(lm3)
```

   + 3d) Fit the model including $age$, $temperature$ and $age^2$. Show the $summary$ and $anova$
output.
```{r}
lm4=lm(length~age+temp+I(age^2),data=fish)
summary(lm4);
anova(lm4)
```

   + 3e)  Show that the coefficient of $age^2$ is the same as that in the regression of 
$e_1$ on $e_2$.  Ans:  The coefficient of $age^2$ equals ??? in both cases.
```{r}
coef(lm4)
coef(lm3)
#answer:according to the data, bothe are -.2564
```

   + 3f) Use {Step 3} in the notes  to show that the coefficients in the $lm$ fit are the
same as those reconstructed from the three stage regression process.  (This is what we did in
class for the tree data.  That is, substitute for $e_1$ and $e_2$ in the equation
$e_1 = \alpha  e_2$, where $\alpha$ is the coefficient from the 3rd regression.
Isolate length on the left hand side, and calculate the regression coefficients on the
right hand side.)
```{r}
coef(lm1)+coef(lm3)*(-coef(lm2))
coef(lm4)
```

   + 3g) Show that the residual sum of squares from the third regression equals that of the $lm$
fit to the full model.  Ans:  The regression SS equals ??? in both cases.
```{r}
anova(lm3)
anova(lm4)
#answer: according to the data both residual sum of squares is 6372687
```

   + 3h) Show that $SSR(\beta_2| \beta_1)$, sometimes written as $S(\beta_2| \beta_1)$,
the extra regression sum of squares explained by
$age^2$ is the same in the third regression as in the $anova$ output for the full model.
Ans:  the regression sum of squares is same in both cases.
```{r}
#according to the above results, tge resgression sum of squares for age square is 8387191 
```

   + 3i) Do the added variable plot for $length$ versus $age^2$ given that the linear effects of
$age$ and $temperature$ have been removed from each variable. 

```{r}
plot(e2,e1,main = "added variable plot for age^2")
```

  Comment on whether it appears appropriate that a quadratic term in
  age be included in the model for length.  (Hint: if this is the case,
  then then this added variable plot should look linear, but not horizontal.)
```{r}
#answer: relationship between e1 and e2 looks close to linear 
```




4.   Data were collected to relate the compressive strength of
concrete ($Y$) to the following set of predictors: 
age ($X_1$), cement ($X_2$), and water ($X_3$).  A number of
regression models were fit, some of which included one or more
quadratic terms  ($X_1^2$, $X_2^2$, $X_3^2$),  and some of which 
included interaction terms ($X_1 X_2, X_1 X_3, X_2 X_3$).  
There were 100 observations in the data set.

  The following table gives the error sums of squares for several
models which were fit. Each model fit contained an intercept term.


\begin{center}
\begin{tabular}{c c l}
Model&Variables in model&SSE \\
1&$X_1$&200\\
2&$X_1,X_2$&190\\
3&$X_1,X_2,X_3$&180\\
4&$X_1,X_2,X_3,X_1^2,X_2^2,X_3^2$&140\\
5&$X_1,X_2,X_3,X_1 X_2, X_1 X_3, X_2 X_3$&150\\
6&$X_1,X_2,X_3,X_1 X_2, X_1 X_3, X_2 X_3, X_1^2,X_2^2,X_3^2$&120\\
\end{tabular}
\end{center}

For each of the following parts, state the full and reduced models which
you are comparing (eg Full: model 6, Reduced: model 1),
give the observed value of the test statistic, and
the numerator and denominator degrees of freedom of the test statistic.
If it is not possible to compare the two models using an F test, state
why not.


   + 4a) Test whether all interaction terms and quadratic terms can be removed from model 6.
```{r}
#Full model:6
#reduce model:3
Fvalue=(((180-120)/6)/(120/(100-10)))
Fvalue
1-pf(Fvalue,6,90)
```
   
   + 4b) Compare two quadratic models, one with and one without interaction terms.
```{r}
#Full model:6
#reduce model:4
Fvalue2=(((140-120)/3)/(120/(100-10)))
Fvalue2
1-pf(Fvalue2,3,90)
```
   
   + 4c) Compare models 2 and 5.
```{r}
#Full model:5
#reduce model:2
Fvalue3=(((190-120)/4)/(120/(100-7)))
Fvalue3
1-pf(Fvalue3,4,93)
```
